---
title: "FAD and Family Focus Project"
output: html_document
---

## Objective

In this article I will walk through the steps I took to clean and visualize data from the Family Assessment Device, a standardized assessments issued to clients and parents of Elements Wilderness Therapy Program to measure family functioning.  The company wanted to know two key things:  
* What did general trends for the FAD survey look like
* Did a client participating in an immersive family therapy session (Family Focus) significantly impact FAD scores on average?


Note: This article does not include cleaning scripts.  These scripts are available on github and will be referenced by name throughout.


```{r setup}
library(tidyverse)
library(patchwork)
library(rstatix)
library(knitr)

# Set wd to parent folder to ensure condidentiality of rds files
knitr::opts_knit$set(root.dir = 'C:/Users/cullo/OneDrive/Desktop/One Cloud/Resume 2022/Github Portfolio/' )

```

## Consent Data  

Information related to whether or not a client consented to thier data being used was recorded on a separate survey.  Thus it was important to merge and clean the consent files in order to join them to each individual survey.

```{r consent}
# Read in rds file with aggregrated consent data from consent_clean_PUBLIC.R
consent <- readRDS("rds/consent master list.rds")
# Rename "clientName" to ensure merge is correct
consent <- consent %>% rename(clientName_con = clientName)
```


## FAD

```{r fad}
# Input FAD dataset with imputed values
# Cleaning file - FAD clean_PUBLIC.R
# Imputation file - imputation_FAD_PUBLIC.R
fad_org <- readRDS("rds/fad_imput_full.rds")


# Merge consent data
fad_con <- merge(fad_org, consent, by = (c("otClientId", "assignedToRelationship")))

# Remove cases that declined consent and extra name column and change score col to fad
fad <- fad_con %>% 
  filter(consent == "Consent") %>% 
  select(-clientName_con) %>% 
  rename(fad = score)

```


```{r fad plot}
# Create table for use in legend
fad_table <- fad %>% filter(instance == "DOA") %>% group_by(surveyID) %>% tally()

# Rename instances for clarity on graph
fad <- fad %>% mutate(instance_rename = case_when(
  instance == "DOA" ~ "Admission",
  instance == "DOD" ~ "Discharge", 
  instance == "MO6" ~ "6 Months Post",
  instance == "YR1" ~ "1 Year Post"))
fad$instance_rename <- factor(fad$instance_rename, levels = c("Admission", "Discharge", "6 Months Post", "1 Year Post"))

# Plot of overall FAD scores
fad.overall.plot <- fad %>% ggplot(aes(color = surveyID)) + 
  stat_summary(aes(x=instance_rename, y=fad), 
               fun = mean, geom = "point") +
  stat_summary(aes(x=instance_rename, y=fad, 
                   group = surveyID), 
               fun = mean, geom = "line") +
  #labs(title = "Overall FAD Scores", subtitle = "How does EWP impact perceptions of family dynamics?")+
  coord_cartesian(ylim = c(1.8,2.35))+
  scale_color_discrete(labels = paste(fad_table$surveyID, fad_table$n, sep = " = "))+
  theme_minimal() +
  geom_hline(aes(yintercept = 2, 
                 linetype = "Clinical Cutoff"
                 )) +
  scale_linetype_manual(name = "",
                        values = c("Clinical Cutoff" = "dashed")) +
  theme(legend.position = c(1,1), legend.justification = c(1,1))+
  xlab("Time") + ylab("FAD Score")+
  labs(title = "Overall FAD Scores",
       subtitle = "Measure of Family Functioning") +
  theme(text = element_text(size = 15)) +
  labs(color = "")

fad.overall.plot

# save as jpg
ggsave("FAD_Outcomes/fig/fad_overall_plot.jpg", fad.overall.plot, width = 10, height = 7)

```

## Family Focus Integration

The client wanted to know how participation in Family Focus impacted the results of these two outcome surveys.  First, I cleaned the data which individual participated in Family Focus (family focus CLEAN_public.R).  Then I joined this data set to both the yoq and fad datasets (shown below).  Next, I produced plots for both measures broken out by Family Focus status.

```{r family focus merge}

# Import family focus (ff) data and master list of names of clients
ff_clients <- readRDS("rds/family focus clients.rds")   # 276

# JOIN TO FAD 
fad <- left_join(fad, ff_clients, by = c("clientName", "otClientId"))
# Edit ff column to be Yes/No
fad[,"ff"][is.na(fad[,"ff"])] <- "No"
fad <- fad[!duplicated(fad[1:13]),]

```

```{r fad by family focus plots}
## table for legend
no.ff.tally <- fad %>% 
  group_by(surveyID, ff) %>% 
  filter(ff == "No") %>% 
  filter(instance == "DOA") %>% tally()
# fad plot of no family focus clients
no.ff <- fad %>% filter(ff == "No") %>% 
  ggplot(aes(color = surveyID)) + 
  stat_summary(aes(x=instance, y=fad), fun = mean, geom = "point") +
  stat_summary(aes(x=instance, y=fad, group = surveyID), fun = mean, geom = "line") +
  geom_abline(intercept = 2, slope = 0, linetype = "dashed") +
  labs(title = "No Family Focus")+
  coord_cartesian(ylim = c(1.75,2.3))+
  scale_color_discrete(labels=paste(no.ff.tally$surveyID, no.ff.tally$n, sep = " = ")) +
  theme_minimal() + 
  theme(legend.position = c(1,1), legend.justification = c(1,1))


# fad table with family focus
yes.ff.tally <- fad %>% 
  group_by(surveyID, ff) %>% 
  filter(ff == "Yes") %>% 
  filter(instance == "DOA") %>% tally()

# fad plot with family focus
yes.ff <- fad %>% filter(ff == "Yes") %>% 
  ggplot(aes(color = surveyID)) + 
  stat_summary(aes(x=instance, y=fad), fun = mean, geom = "point") +
  stat_summary(aes(x=instance, y=fad, group = surveyID), fun = mean, geom = "line") +
  geom_abline(intercept = 2, slope = 0, linetype = "dashed") +
  labs(title = "With Family Focus")+
  coord_cartesian(ylim = c(1.75,2.3))+
  scale_color_discrete(labels=paste(yes.ff.tally$surveyID, yes.ff.tally$n, sep = " = ")) +
  theme_minimal() +
  theme(legend.position = c(1,1), legend.justification = c(1,1))

# Family Assessment Device, family focus vs none plots, 
ff.dif <- no.ff + yes.ff
ff.dif

ggsave("FAD_Outcomes/fig/fad_by_family_focus.jpg", ff.dif, width = 10, height = 7)


```


```{r fad by family focus plots}
# Check for statistical difference between ff and no ff
fad_ttest <- fad %>% 
  group_by(instance, surveyID) %>% 
  t_test(fad ~ ff) %>%
  adjust_pvalue(method = "BH") %>% 
  add_significance()

```

Participation in a family focus was not shown to impact the scores of individuals at any time point in a statistically significant way.  







